---
title: "projekt"
output: html_document
date: "2022-12-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


###   GETTING STARTED WITH LEAFLET

# Try to work through down this script, observing what happens in the plotting pane.

# Review favorite backgrounds in:
# https://leaflet-extras.github.io/leaflet-providers/preview/
# beware that some need extra options specified

# To install Leaflet package, run this command at your R prompt:
## We will also need this widget to make pretty maps:
```{r}
#install.packages("leaflet")
#install.packages("htmlwidgets")
```

# Activate the libraries
```{r}
library(leaflet)
library(htmlwidgets)
library(tidyverse)
```

########## Example with Markers on a map of Europe


# First, create labels for your points
```{r}
popup = c("Robin", "Jakub", "Jannes")
```

# You create a Leaflet map with these basic steps: you need to run the whole chain of course
```{r}
leaflet() %>%                                 # create a map widget by calling the library
  addProviderTiles("Esri.WorldPhysical") %>%  # add Esri World Physical map tiles
  addAwesomeMarkers(lng = c(-3, 23, 11),      # add layers, specified with longitude for 3 points
                    lat = c(52, 53, 49),      # and latitude for 3 points
                    popup = popup)            # specify labels, which will appear if you click on the point in the map

```

### Let's look at Sydney with setView() function in Leaflet
```{r}
leaflet() %>%
  addTiles() %>%                              # add default OpenStreetMap map tiles
  addProviderTiles("Esri.WorldImagery",       # add custom Esri World Physical map tiles
                   options = providerTileOptions(opacity=0.5)) %>%     # make the Esri tile transparent
  setView(lng = 151.005006, lat = -33.9767231, zoom = 10)              # set the location of the map 
```



# Now let's refocus on Europe again
```{r}
#leaflet() %>% 
 # addTiles() %>% 
  #setView( lng = 2.34, lat = 48.85, zoom = 5 ) %>%  # let's use setView to navigate to our area
  #addProviderTiles("Esri.WorldPhysical", group = "Physical") %>% 
  #addProviderTiles("Esri.WorldImagery", group = "Aerial") %>% 
  #addProviderTiles("MtbMap", group = "Geo") %>% 
```


```{r}
#addLayersControl(                                 # we are adding layers control to the maps
 # baseGroups = c("Geo","Aerial", "Physical"),
  #options = layersControlOptions(collapsed = T))
```

# click the box in topright corner in your Viewer 
# to select between different background layers


########## SYDNEY HARBOUR DISPLAY WITH LAYERS
# Let's create a more complicated map 

# Set the location and zoom level
```{r}
leaflet() %>% 
  setView(151.2339084, -33.85089, zoom = 13) %>%
  addTiles()  # checking I am in the right area
```


# Bring in a choice of esri background layers  

# Create a basic basemap
```{r}
l_aus <- leaflet() %>%   # assign the base location to an object
  setView(151.2339084, -33.85089, zoom = 13)
```


# Now, prepare to select backgrounds
```{r}
esri <- grep("^Esri", providers, value = TRUE)
```


# Select backgrounds from among provider tiles. To view them the options, 
# go to https://leaflet-extras.github.io/leaflet-providers/preview/
```{r}
for (provider in esri) {
  l_aus <- l_aus %>% addProviderTiles(provider, group = provider)
}
```

```{r}
l_aus
```

### Map of Sydney, NSW, Australia
# We make a layered map out of the components above and write it to 
# an object called AUSmap
```{r}
AUSmap <- l_aus %>%
  addLayersControl(baseGroups = names(esri),
                   options = layersControlOptions(collapsed = FALSE)) %>%
  addMiniMap(tiles = esri[[1]], toggleDisplay = TRUE,
             position = "bottomright") %>%
  addMeasure(
    position = "bottomleft",
    primaryLengthUnit = "meters",
    primaryAreaUnit = "sqmeters",
    activeColor = "#3D535D",
    completedColor = "#7D4479") %>% 
  htmlwidgets::onRender("
                        function(el, x) {
                        var myMap = this;
                        myMap.on('baselayerchange',
                        function (e) {
                        myMap.minimap.changeLayer(L.tileLayer.provider(e.name));
                        })
                        }") %>% 
addControl("", position = "topright")
```

# run this to see your product
```{r}
AUSmap
```

########## SAVE THE FINAL PRODUCT

# Save map as a html document (optional, replacement of pushing the export button)
# only works in root

```{r}
saveWidget(AUSmap, "AUSmap.html", selfcontained = TRUE)
```


###################################  YOUR TASK NUMBER ONE


# Task 1: Create a Danish equivalent of AUSmap with esri layers, 
# but call it DANmap

```{r}
Danmap <- leaflet() %>%   # assign the base location to an object
  setView(9.669434, 56.104165, zoom = 13)
```


# Husk at køre Esri hvis du laver et kort i dit eget 

```{r}
for (provider in esri) {
  Danmap <- Danmap %>% addProviderTiles(provider, group = provider)
}
```

```{r}
Danmap
```

```{r}
DANmap <- Danmap %>%
  addLayersControl(baseGroups = names(esri),
                   options = layersControlOptions(collapsed = FALSE)) %>%
  addMiniMap(tiles = esri[[1]], toggleDisplay = TRUE,
             position = "bottomright") %>%
  addMeasure(
    position = "bottomleft",
    primaryLengthUnit = "meters",
    primaryAreaUnit = "sqmeters",
    activeColor = "#3D535D",
    completedColor = "#7D4479") %>% 
  htmlwidgets::onRender("
                        function(el, x) {
                        var myMap = this;
                        myMap.on('baselayerchange',
                        function (e) {
                        myMap.minimap.changeLayer(L.tileLayer.provider(e.name));
                        })
                        }") %>% 
  addControl("", position = "topright")
```

```{r}
DANmap
```

# Cordinates to himmelbjerget: 56.104165,9.669434

########## ADD DATA TO LEAFLET

# In this section you will manually create machine-readable spatial
# data from GoogleMaps: 

### First, go to https://bit.ly/CreateCoordinates1
### Enter the coordinates of your favorite leisure places in Denmark 
      # extracting them from the URL in googlemaps, adding name and type of monument.
      # Remember to copy the coordinates as a string, as just two decimal numbers separated by comma. 

# Caveats: Do NOT edit the grey columns! They populate automatically!

### Second, read the sheet into R. You will need gmail login information. 
      # watch the console, it may ask you to authenticate or put in the number 
      # that corresponds to the account you wish to use.

```{r}
#install.packages("googlesheets4")
```

# Libraries
```{r}
library(tidyverse)
library(googlesheets4)
library(leaflet)
gs4_deauth()
```

### THe gs4_deauth() function is used to help load the dataset from googledocs, because it shoved an error in the try at downloading. It is the same as adding range = "SA2022" to the below code string.

# Read in a Google sheet

```{r}
places <- read_sheet("https://docs.google.com/spreadsheets/d/1PlxsPElZML8LZKyXbqdAYeQCDIvDps2McZx1cTVWSzI/edit#gid=124710918",
                     col_types = "cccnncnc", range = "SA2022")
glimpse(places)
```

## These below functions were used to try and discover the problem with downloading the dataset from googledocs.
```{r}
#rlang::last_error()
#rlang::last_trace()
```

# load the coordinates in the map and check: are any points missing? Why?
```{r}
leaflet() %>% 
  addTiles() %>% 
  addMarkers(lng = places$Longitude, 
             lat = places$Latitude,
             popup = places$Description, 
             clusterOptions = markerClusterOptions())
```


#########################################################

#This is for adding more information to the popups
```{r}
leaflet() %>% 
  addTiles() %>% 
  addMarkers(lng = places$Longitude, 
             lat = places$Latitude,
             popup = paste(places$Placename,'<br>',  places$Type,'<br>', places$Stars1_5,'<br>', places$Description),
             clusterOptions = markerClusterOptions())
```

##You can bouth do '<br>' for a brake, and you can also do a comma "," depending how you would like to have the information put up.

# Task 2: Read in the googlesheet data you and your colleagues, populated with data into the DANmap object you created in Task 1.
## This task was solved by simply downloading the googledoc dataset we had created in the class. If you go to the dataset now, there might be more points now, because people were still adding more points when we got to that point in the exercise. 

# Task 3: Can you cluster the points in Leaflet? Google "clustering options in Leaflet"
## You need to add clusterOptions = markerClusterOptions() to your addMarkers() function, this makes it so if there are multiple points near each other you can get the map to cluster them together so they are easy to navigate.

# Task 4: Look at the map and consider what it is good for and what not.
## After we changed the name to Placenames instead of Description it works really good. Makes it easy to find the individual places. At the same time making it easy to navigate in your map, it might be a bit annoying if you by accident press one of the points. 

# Task 5: Find out how to display notes and classifications  in the map.
## To solve this we can use the paste() funktion after popup=, so it is written like popup=paste(). And as can be seen in the map there are now 4 different informations, for the different recommendations made from the team. 

