---
title: "Map1"
output: html_document
date: "2022-12-09"
---

#Pakages
```{r}
library(tidyverse)
library(leaflet)
library(mapproj)
library(maps)
library(dplyr)
library(crosstalk)
```

# Dataset: https://www.kaggle.com/datasets/rtatman/historical-american-lynching

#Loading the dataset, where Simone Broni had used a lot of time to add latitude and longtitude but decided to do another project. So credit is given to her to add this to the dataset.
```{r}
Lynching_data <- read_csv("data/Lynching-Data-Simone-Fixed-xlsx (1).csv")
States <- map_data("state")
```

#The first problem found is that Latitude for some reason was a charcter value instead of numeric. So to fix it I used this peace of code:
### This code was inspired by https://stackoverflow.com/questions/2288485/how-to-convert-a-data-frame-column-to-numeric-type
```{r}
Lynching_data <- transform(Lynching_data, Latitude = as.numeric(Latitude))
```

#Tried this code to make the Latitude numeric and it did not work.
```{r}
Lynching_data$Latitude <- as.numeric(Lynching_data$Latitude) 
Lynching_data$Longitude <- as.numeric(Lynching_data$Longitude)
```

#Hereafter I used the map exercise thought to us in class, to create the following map with clusters and added information to these clusters.
##In the clusters I chose to add the information: Statem Year, Offense, Race and Victim(The name if known)
```{r}
esri <- grep("^Esri", providers, value = TRUE)
```


```{r}
Lynching_map <- 
  leaflet() %>% 
  addTiles() %>% 
  addMarkers(lng = Lynching_data$Longitude, 
             lat = Lynching_data$Latitude,
             popup = paste(Lynching_data$State,'<br>',  Lynching_data$Year,'<br>', Lynching_data$Offense,'<br>', Lynching_data$Race,'<br>',Lynching_data$Victim),
             clusterOptions = markerClusterOptions())
```
```{r}
Lynching_map
```
#This map can now easly be navigated in, and you can get the individual lynching victims information. There is the option to add more information, but I did not se the relevans in doing so. 
##As will be discussed in the exam project the map also shows the imbalance of the dataset.


#I then decided to try making a map with a slider.
## The slider could  have helped with an analysis for specifik timeperiods, but there was sadly not enough space due to the charecter limit.  Here I got help from Asmus, that had found a solution from this stackocerflow: https://stackoverflow.com/questions/62849300/r-leaflet-add-a-range-slider-to-filter-markers-without-shiny
##I would also have trie to add a couter of the datapoints, that would change with how you moved the slider.

```{r maak shared data object}
shared_data <- SharedData$new(Lynching_data)
```

```{r genereer output}
filter_slider("Lynching_data$Year", "Date", shared_data, ~Lynching_data$Year, width = "100%")
leaflet(shared_data, width = "100%", height = 800) %>%
  leaflet::addTiles() %>%
  leaflet::addMarkers() 
```

##The slider will only show up in the knitted htlm file.


